---
title: "Regression_Tutorial"
author: "Daniel T. Citron"
date: "8/13/2018"
output: html_document
---
# Logistic Regression Tutorial
In which I finally learn how to implement logistic regression - this will include single variable binomial logit model; multivariable binomial logit model; multinomial logit model.

Logit model and logistic regression can be used interchangeably(?)

```{r Read in Libraries}
library(ggplot2)
library(foreign)
library(nnet)
library(reshape2)
library(data.table)
```

## Single Variable
Tutorial data taken from Wikipedia: https://en.wikipedia.org/wiki/Logistic_regression

Can we predict how many hours 

```{r Logistic Regression with Single Covariate}
passing.test.scores = data.table(
  hours = c(0.50,	0.75,	1.00,	1.25,	1.50,	1.75,	1.75,	2.00,	2.25,	2.50,	2.75,	3.00,	3.25,	3.50,	4.00,	4.25,	4.50,	4.75,	5.00,	5.5),
  pass = c(0, 0,	0, 0,	0, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 1, 1, 1, 1, 1)
  )
mylogit <- glm(pass ~ hours, data = passing.test.scores, family = "binomial")
summary(mylogit)
```

Since we have one dependent variable, we can make a plot:
```{r Plot Probability vs. Hours studied}
x <- seq(0,6,.1)
plot(passing.test.scores$hours, passing.test.scores$pass)
lines(x, 1/(1 + exp(-(mylogit$coefficients[[1]] + mylogit$coefficients[[2]]*x))))
```


## Multiple Variable
Tutorial data taken from UCLA: https://stats.idre.ucla.edu/r/dae/logit-regression/

This data set has a binary response (`admit`) that says whether or not someone got admitted.  There are three dependent covariates: `gpa`, `gre` score, and `rank`.  `gpa` and `gre` will be treated as continuous, but the `rank` variable will be treated as discrete.
```{r Logistic Regression with Multiple Covariates}
mydata <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")
summary(mydata)
```

Logistic regression (logit model) is one option, but the probit model is also an option - it's a similarly-shaped log-odds function.

We first need to convert the `rank` variable to a `factor` to reflect its categorical status:
```{r Using a Logit Model}
mydata$rank <- factor(mydata$rank)
mylogit <- glm(admit ~ gre + gpa + rank, family = "binomial", data = mydata)
summary(mylogit)
```

From the output, we can see that lower rank decreases one's probability of admission. GPA is a positive predictor of admission, and GRE score is a very weakly positive predictor of admission.

Use the `confint` function to obtain confidence intervals for the coefficient estimates.
```{r}
# using profiled log-likelihood
confint(mylogit)
# using Standard Errors
confint.default(mylogit)
```

```{r}
AIC(mylogit)
```

One small problem: 
