---
title: "Ecological Models and Data in R"
output: html_notebook
---
```{r include = FALSE}
install.packages("numDeriv")
install.packages("bbmle")
library(stats4)
library(bbmle)
```

#Chapter 5

##5.2.1 Simple examples

First example: fit a linear model with normally-distributed errors
```{r}
x = 1:20
a = 2
b = 1
```

Deterministic part of the model
```{r}
y_det = a*x + b
```

Add noise to the model - pick 20 random, normally-distribute deviates with mean equal to the deterministic equation and standard deviation = 2
```{r}
# Simulate, by generating a noisy data set
y_noise = rnorm(20, mean = y_det, sd = 2)
# Alternatively (but less generalizably),
# y_noise = y_det + rnorm(20, sd = 2)
```

Fit the linear model
```{r}
lm(y_noise~x)
```


Second example: fit a hyperbolic model with negative binomially-distributed errors
```{r}
a = 20
b = 1
k = 5
```

Generate predictor variable x:
```{r}
x = runif(50, min = 0, max = 5)
```

Generate deterministic mean and the noisy data
```{r}
y_det = a*b/(b + x)
# Simulate, by generating a noisy data set
y = rnbinom(50, mu = y_det, size = k)
```

Fit the model
```{r}
m = mle2(y_det ~ dnbinom(mu = a * b * x/(b + x), size = k), start = list(a = 15, b = 1, k = 5))
```

## 5.3 Power Analysis
Finding out the (frequentist) statistical power, the probability of failing to reject the null hypothesis when it is false.

"How do the quality and quantity of my data and the true properties (parameters) of my ecological system affect the quality of my answers to my questions about ecological systems?"

Estimate the statistical power of detecting a linear trend as a function of sample size.  Simulate to determine whether it is likely that we can reject the null hypothesis in a single "experiment."
```{r}
N = 10
x = seq(1,20,20./N)
a = 2
b = 1
y_det = a + b * x
y_noise = rnorm(N, mean = y_det, sd = 2)
m = lm(y_noise ~ x)
coef(summary(m))["x", "Pr(>|t|)"]
```

Repeat 400 times and see how many times we can reject our hypothesis:
```{r}
nsim = 400
pval = numeric(nsim)
for (i in 1:nsim){
  y_det = a*x + b
  y_noise = rnorm(N, mean = y_det, sd = 2)
  m = lm(y_noise ~ x)
  pval[i] = coef(summary(m))["x", "Pr(>|t|)"]
}
```
And calculate the power:
```{r}
sum(pval < 0.05)/nsim
```

